Parallel global optimization on GPU \thanks{The revised version of the paper was supported by the Russian Science Foundation, project No 15-11-30022 ``Global optimization, supercomputing computations, and applications''}

### 2019-10-07 2016/JoGO_GPU/BarkalovGergel.tex
JOGO2016 paper final version

This paper considers the problems of Lipschitzian global optimization that are frequently encountered in applications. They are characterized by a considerable time required to calculate the function value at one point (in comparison with the processing of the results of such calculation). For example, the objective function can be defined by solving SLAE, ODE, PDE, or Monte Carlo simulation. Currently, GPU may be used to solve all these types of problems \cite{RefKindratenko}. Therefore, calculation of values of an optimized function can be done using a GPU, and the main objective of the algorithm (working at CPU) is to efficiently select points of parallel trials. The article contains the results of an investigation into the GPU implementation of the global search parallel algorithm developed according to the information-statistical approach presented in the monograph \cite{RefStrongin2000}. Some preliminary results of research in this area have been published in \cite{RefLebedevBarkalov}.
%new text end

@@

\begin{figure}
\begin{minipage}{0.32\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig1a.JPG} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}{0.32\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (b)}
\end{minipage}
\begin{minipage}{0.32\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig1c.JPG} \\ (c)}
\end{minipage}
\caption{Elvolvents in two dimensions with (a) $m=3$, (b) $m=4$ and (c) $m=5$}
\label{fig:0}
\end{figure}
%new text end

@@

@@

Let us use a parallel algorithm of global search for solving a problem from section \ref{sec:2} with the numbers of parallel trials $p=2$ and $p=4$. With $p=2$ the number of iterations was $569$, and the number of trials was $1138$. With $p=4$ the number of iterations was reduced to $270$, and $1080$ trials were carried out. Fig.~\ref{fig:2} shows level lines of the function with points of the trials performed by the method with $p=2$. Points of redundant trials carried out using a parallel algorithm are marked in the figure by "+". The concept of redundancy is discussed in more detail in the following section.
\begin{figure}
	\center
  \includegraphics[width=0.75\textwidth]{fig3.jpg} 
  \caption{Minimization of a function of two variables by the parallel algorithm}
  \label{fig:2}
\end{figure}

@@

It should be noted that widely known test problems from the field of multidimensional global optimization are characterized by small time required to calculate the objective function value. Usually, such a calculation is nothing else than a summation of several (according to problem dimension) values of elementary functions. For example, the well known Rastrigin test function, the global minimum of which is located in $y^\ast = (0,\dots,0)$ with $\varphi(y^\ast)=0$, is set by the expression
\begin{equation}\label{eq:33}
\varphi(y)=10N+\sum^{N}_{i=1}\left[y^2_i-10\cos(2\pi y_i)\right], \; -2.2\leq y_i \leq 1.8, \; 1\leq i \leq N.
\end{equation}

@@

The results of our experiments show some obvious advantages of the GPU algorithm in comparison with the CPU algorithm. The algorithm, which uses the maximal number of cores of CPU, works slower than any of the given GPU implementations. In addition, the results show a considerable speedup in the number of iterations. For example, on average only 353 parallel iterations on GPU were sufficient to solve problems of the most difficult class considered -- 5-\textit{Hard}, whereas even with the use of all computing CPU cores over eleven thousand iterations were required. Finally, the GPU algorithm is almost non-redundant in relation to the CPU algorithm.

@@

\end{document}
% end of file template.tex

>>>>

This paper considers the problems of Lipschitzian global optimization that are frequently encountered in applications. They are characterized by a considerable time required to calculate the function value at one point (in comparison with the processing of the results of such calculation). For example, the objective function can be defined by solving SLAE, ODE, PDE, or Monte Carlo simulation. Currently, GPU can be used to solve all these types of problems. 
%rev4 text begin
Moreover, a set of problems can be solved on GPU simultaneously. For example, algorithm of solving of a large amount of independent SLAE on GPU is considered in \cite{RefKindratenko}, chapter 5. Results of simultaneously solving $2 \cdot 10^4$ systems of variable size are presented. Algorithm of parallel integration of large numbers of independent ODE systems on GPU is discussed in \cite{RefKindratenko}, chapter 8. Experimental results that show good speedup in solving a set (from $10^3$ up to $10^5$) of independent ODE systems are presented. Assuming that the objective function in global optimization problem is given as SLAE or ODE system many function values can be computed simultaneously.
%rev4 text end

Therefore, calculation of an objective function values can be done using a GPU, and the main goal of the algorithm	 (working at CPU) is to efficiently select points of parallel trials. The article contains the results of an investigation into the GPU implementation of the global search parallel algorithm developed according to the information-statistical approach presented in the monograph \cite{RefStrongin2000}. Some preliminary results of research in this area have been published in \cite{RefLebedevBarkalov}.
%new text end

@@

\begin{figure}[ht]
\begin{minipage}{0.32\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig1a.JPG} \\ (a)}
\end{minipage}
\begin{minipage}{0.32\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig1b.JPG} \\ (b)}
\end{minipage}
\begin{minipage}{0.32\linewidth}
\center{\includegraphics[width=1.0\linewidth]{fig1c.JPG} \\ (c)}
\end{minipage}
\caption{Evolvents in two dimensions with (a) $m=3$, (b) $m=4$ and (c) $m=5$}
\label{fig:0}
\end{figure}
%new text end

@@

\begin{figure}
	\center
  \includegraphics[width=0.75\textwidth]{fig3.jpg} 
  \caption{Minimization of a function of two variables by the parallel algorithm}
  \label{fig:2}
\end{figure}

@@

Let us use a parallel algorithm of global search for solving a problem from section \ref{sec:2} with the numbers of parallel trials $p=2$ and $p=4$. With $p=2$ the number of iterations was $569$, and the number of trials was $1138$. With $p=4$ the number of iterations was reduced to $270$, and $1080$ trials were carried out. Fig.~\ref{fig:2} shows level lines of the function with points of the trials performed by the method with $p=2$. Points of redundant trials carried out using a parallel algorithm are marked in the figure by "+". The concept of redundancy is discussed in more detail in the following section.

@@

It should be noted that widely known test problems from the field of multidimensional global optimization are characterized by small time required to calculate the objective function value. Usually, such a calculation is nothing else than a summation of several (according to problem dimension) values of elementary functions. For example, the well known Rastrigin test function 
\begin{equation}\label{eq:33}
\varphi(y)=10N+\sum^{N}_{i=1}\left[y^2_i-10\cos(2\pi y_i)\right], \; -2.2\leq y_i \leq 1.8, \; 1\leq i \leq N,
\end{equation}
has the global minimum value $\varphi(y^\ast)=0$ at the point $y^\ast = (0,\dots,0)$.

@@

The results of our experiments show some obvious advantages of the GPU algorithm in comparison with the CPU algorithm. The algorithm, which uses the maximal number of cores of CPU, works slower than any of the given GPU implementations. In addition, the results show a considerable decrease of the number of iterations. For example, on average only 353 parallel iterations on GPU were sufficient to solve problems of the most difficult class considered -- 5-\textit{Hard}, whereas even with the use of all computing CPU cores over eleven thousand iterations were required. Finally, the GPU algorithm is almost non-redundant in relation to the CPU algorithm.

@@

\end{document}
